2D Subquadratic Separable Distance Transformation for Path-Based Norms\thanks{This work has been mainly funded by ANR-11-BS02-009 and ANR-11-IDEX-0007-02 PALSE/2013/21 research grants.}

### 2014-02-18 latex/dgci2014/article.tex
ok

\documentclass{llncs}
\usepackage{amsmath,amssymb}
\usepackage{url}
\usepackage{overpic}
\usepackage{enumerate}
\usepackage{graphicx}        % standard LaTeX graphics tool
\usepackage{tikz}        % standard LaTeX graphics tool
\usepackage{subfigure}                                 % authors: subfigures
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}   % authors: last version of algorithm display
\usepackage{todonotes}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}

@@

\title{Generic Separable Algorithms for Distance Transformation: A
  Subquadratic Approach for Path Based Norms\thanks{This work has been
    mainly funded by ANR-11-BS02-009 and ANR-11-IDEX-0007-02
    PALSE/2013/21 research grants.}}

@@

\begin{abstract}\sloppy
In many applications, separable algorithms have demonstrated their
efficiency to perform high performance volumetric computations, such
as distance transformation or medial axis extraction. In the
literature, several authors have discussed about the conditions on the
metric to be considered in a separable approach.  In this article, we
present generic separable algorithms to efficiently compute Voronoi
maps and distance transformations for a large class of
metrics. Focusing to path based norms (chamfer masks, neighborhood
sequences, ...), we detail a subquadratic algorithm to compute such
volumetric transformations in dimension 2. More precisely, we describe
a $O(\log^2{m}\cdot N^2)$ algorithm in dimension 2 for shapes in
$N\times N$ domain with chamfer norm with size $m$.

\keywords{Digital Geometry, Distance Transformation, Chamfer Norms}
\end{abstract}

@@

Since early works on digital geometry, distance transformation plays
an important role in many applications
\cite{Rosenfeld1966,Rosenfeld1968}. Given a finite input shape
$X\subset \Z^n$, the distance transformation labels each point in $X$
with the distance to the closest point in $\Z^n \setminus X$. Since
such characterization is parametrized by a distance function, many
works have been done to address this distance transformation problem
with a trade-off between algorithmic performances and the
\emph{accuracy} of the digital distance function with respect to the
Euclidean one.  Hence, authors have considered distances based on
chamfer masks
\cite{Rosenfeld1968,borgefors,Thiel_IWCIA7,fouard:ivc:2005} or
sequences of chamfer masks
\cite{ROSEN_66,mukherjee,Nagy05,Strand2008,DBLP:conf/dgci/NormandSE13};
the vector displacement based Euclidean distance
\cite{danielson,ragnemalm,MULL_92,Cuisenaire1999_268}; the Voronoi
diagram based Euclidean distance
\cite{BreuEtAl95,maurer_pami} or the square of the
Euclidean distance \cite{SaitTori:94,Hirata,roerdnik}.

\paragraph{Contributions} In this article,

\begin{table}
  \caption{Complexity summary.}
  \label{tab:final}
  \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      Metric &\textsc{Closest}& \textsc{HiddenBy} & Sep. Voronoi Map & Reference\\
      \hline
      $L_2$ & $O(1)$ &  $O(1)$ & $\Theta(n\cdot N^n)$ & \cite{Hirata1996}\\
      $L_\infty$ & $O(1)$ & $O(1)$ &  $\Theta(n\cdot N^n)$ & \cite{roerdnik}\\
      $L_1$ & $O(1)$ &   $O(1)$ &$\Theta(n\cdot N^n)$& \cite{roerdnik}\\
      $L_p$  (exact pred.) & $O(\log{p})$ &
      $O(\log{p}\cdot\log{N})$ & $O(n\cdot
      N^n\cdot\log{p}\cdot\log{N})$& here and  \cite{dgtal}\\
      $L_p$  (inexact pred.) & $O(1)$ &
      $O(\log{N})$ & $O(n\cdot
      N^n\cdot\log{N})$& here and  \cite{dgtal}\\
      2D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
      &$O(\log^2{m}\cdot N^2)$& Theorem \ref{them}\\
  %    3D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
  %    &$O(\log^2{m}\cdot N^3)$& Theorem \ref{them2}\\
  %    nD Chamfer norm &  $O(n\cdot\log{f})$ &$O(n\cdot log^2{f}\cdot\log{N})$
  %    & $O(n^2\cdot N^n\cdot \log{f}\cdot\log{N})$& Corollary \ref{coro2}\\
      2D Neig. sequence norm  & $O(\log{f})$&
      $O(\log^2{f})$& $O(\log^2{}f\cdot N^2)$& Th.~\ref{them} and  \cite{DBLP:conf/dgci/NormandSE13}\\
      \hline
    \end{tabular}
  \end{center}
\end{table}

\section{Preliminaries}
\label{sec:preliminaries}
Let us first define some elementary concepts and objects.
\begin{definition}[Metrics and Metric Space]
  \label{def:distance}
  A metric is a map $d$ from $E$ to a sub-group $F$ of $\R$ such that
  $\forall a,b,c\in E$,
  \begin{align}
    (\text{non-negative})\quad & d(a,b)\geq 0\\
    (\text{identity of indiscernibles}) \quad&  d(a,b)= 0
    \Leftrightarrow a=b\\
    (\text{symmetry})\quad &  d(a,b)=d(b,a)\\
    (\text{triangular inequality})\quad &   d(a,c) \leq d(a,b) + d(b,c)
  \end{align}
$(E, F, d)$ is called a metric space.
\end{definition}
If property $(4)$ is omitted, we are defining a distance function
instead of a metric. Note that in the following, we will only consider
metric spaces with properties $(1-4)$. Furthermore, we
consider  metrics induced by a vector space norm:
\begin{definition}[Norm and metric induced by a norm]
  \label{def:distance}
  Given a vector space EV, a norm is a map $g$ from  $EV$ to a sub-group
  $F$ of $\R$ such that $\forall \vec{x},\vec{y}\in EV$,
  \begin{align}
    (\text{non-negative})\quad & g(\vec{x})\geq 0\\
    (\text{identity of indiscernibles}) \quad&  g(\vec{x})= 0
    \Leftrightarrow \vec{x}=\vec{0}\\
    (\text{triangular inequality})\quad &   g(\vec{x}+\vec{y}) \leq
    g(\vec{x}) + g(\vec{y})\\
    (\text{homogeneity})\quad &  \forall \lambda\in\R, \quad
    g(\lambda\cdot\vec{x}) = |\lambda|\cdot g(\lambda\cdot\vec{x})
  \end{align}
$d(a,b) \EqDef g(b-a)$ is the metric induced by the
  norm $g$.
\end{definition}
Note that the above definition can be extended from vector spaces to
\emph{modules} on a commutative ring ($\Z^n$ being a module on $\Z$
but not a vector space) \cite{Thiel_hdr}.
From the literature, path-based approaches (Chamfer masks, -weighted-
neighborhood sequences...)  aim at defining metrics induced by norms
in metric spaces $(\Z^n, \Z , d)$.  Note that (weighed, with $w_i\geq 0$) $L_p$ norms inducing
metrics
\begin{equation}
    d_{L_p} (a,b) = \left ( \sum_{i=0}^n w_i|a_i-b_i |^p \right )^{\frac{1}{p}}\;,
  \end{equation}
characterize metric spaces $(\Z^n, \R , d_{L_p})$. However,
rounding the distance function $(\Z^n,\Z, \lceil d_{L_p}\rceil)$ implies a
metric space \cite{klette_book}.

\begin{definition}[Distance Transformation and Voronoi Map]
  The distance transform $DT_X$ associated with a digital metric space
  $(\Z^n,\Z,d)$ is a map  $X\rightarrow\Z$ such that
  \begin{equation}
    DT_X(a) = \min_{b\in \Z^n\setminus X} \{d(a,b)\}\;.
  \end{equation}
The Voronoi map is the map $X\rightarrow\Z^n$:
  \begin{equation}
    \Pi_X(a) = \argmin_{b\in \Z^n\setminus X} \{d(a,b)\}\;.
  \end{equation}
\end{definition}
Voronoi map $\Pi_X$ corresponds to the intersection between the
continuous Voronoi diagram for the metric $d$ of points
$\Z^n\setminus$ and the lattice $\Z^n$. If a digital point $a$ belongs
to a Voronoi diagram $d-$facet ($1\leq d< n$), $a$ is equidistant to 2
or more points in $\Z^n\setminus X$ but only one is considered in
$\Pi_X(a)$ (please refer to\cite{Couprie2007} for complete digital
Voronoi map construction).

\begin{definition}[Chamfer Mask]
  A weighted vector is a pair $(\vec{v},w)$ with $\vec{v}\in\Z^n$ and
  $w\in\N^*$. A chamfer mask $\M$ is a central-symmetric set of weighted
  vectors with non-empty vectors and containing at least a basis of
  $\Z^n$.
\end{definition}

In the literature
\cite{borgefors,Thiel_hdr,Strand2008,normandHDR,fouard:ivc:2005} many
authors have proposed algorithmic and/or analytic approaches to
construct Chamfer masks approximating the Euclidean metric and being
induced by norm. In the following, we focus on such \emph{chamfer
  norms}.  To evaluate distances between two digital points for a
given chamfer metric, direct formulations have been proposed with
simple geometrical interpretation:
\begin{definition}[Rational ball, minimal H-representation \cite{Thiel_hdr,DBLP:journals/pr/NormandE09}]
  Given a Chamfer mask $\M$ inducing a norm, the rational ball
  associated with $\M$ is the polytope
  \begin{equation}
\label{eq:4}
    \B = conv\left\{ \frac{\vec{v}_i}{w_i};\; (\vec{v}_i,w_i)\in\M \right \}\,.
  \end{equation}
Rational ball $\B$ can also be described in a H-representation of
polytope with minimal parameter \cite{DBLP:conf/dgci/NormandSE13}:
\begin{equation}
  P = \{ x\in\Z^n; Ax \leq y\;\}\quad \text{ such that }  \forall
  i\in[1\ldots f], \exists x\in P\quad A_ix=y_i\,.\footnote{$A_i$
    being the $i^{th}$ row of $A$.}
\end{equation}
$f$ being the number of row in $A$ and the number of facets in $\B$,   and is thus related to $|\M|$.
\end{definition}
An important result for distance computation can be formalized as follows:
\begin{proposition}[Direct Distance Computation \cite{DBLP:journals/pr/NormandE09}]
  Given a chamfer mask $\M$ inducing a norm and $(A,y)$ its minimal
  parameter H-representation, then for any $a\in\Z^n$, the chamfer
  distance of $a$ from the origin is
  \begin{equation}
    \label{eq:1}
    d_\M(O,a) =  \max_{1\leq i \leq l} \{ A_i a^T\}\,.
  \end{equation}
\end{proposition}

Among path-based digital metric, we can also introduce (weighted)
neighborhood sequences
\cite{ROSEN_66,mukherjee,Nagy05,Strand2008,DBLP:conf/dgci/NormandSE13}
whose objective is to have better approximation of the Euclidean
metric from sequences of elementary chamfer masks.  A key result have
been demonstrated in \cite{DBLP:journals/pr/NormandE09} stating
that for such distance functions, a minimal parameter polytope
representation exists and that distances can be from similar
expression to (\ref{eq:1}):
\begin{equation}
  d(O,a)  =\max_{1\leq i \leq l} \{ f_i(A_i a^T)\}\,,
\end{equation}
($f_i$ being and integer sequence characterizing the neighborhood
sequence metric).

 In the following and for the sake of simplicity, we describe our
 algorithms focusing to chamfer norms but similar results can be
 obtained for more generic path-based metric such neighborhood
 sequences.
\section{Separable Distance Transformation}
\label{sec:separ-dist-transf}
\subsection{Voronoi Map from Separable Approach and Metric Conditions}
\label{sec:voronoi-map-from}
In \cite{Hirata1996,SaitTori:94,Breu1995,Meijster2000,Maurer2003},
several authors have described optimal in time and separable
techniques to compute error-free Voronoi maps or distance
transformations for $L_2$ and $L_p$ metrics. Separability means that
computations are performed dimension by dimension, allowing fast
computations and multithread implementations (see
\cite{dcoeurjo_ChapDT} for an overview).  In the following, we
consider Voronoi map approach as defined in \cite{Breu1995} (see
Fig.~\ref{fig:2Dbreu}): Let us first define hyper-rectangular domain
$I_X: [0..N_1]\times\ldots[0..N_n] \rightarrow \{0,1\}$ such that
$X\subset I_X$ and $I_X(a)= 1$ for $a\in [0..N_1]\times\ldots[0..N_n]$
iff $a\in X$ ($I_X(a)=0$ otherwise). As illustrated in
Fig.~\ref{fig:2Dbreu} in dimension 2, we first process each row of the
input image to create independent 1D Voronoi maps along the first
dimension for the metric. Then, for each further dimension, the
partial Voronoi map $\Pi_X$ is updated using one dimensional
independent processes on image spans along the $i^{th}$
dimension.

%% \begin{figure}
%%   \begin{center}
%%      \begin{overpic}[draft=false,width=10cm]{figs/edt_dcoeurjo}
%%        \put(30,5){$Candidates_S$}
%%        \put(57,0){$S$}
%%      \end{overpic}
%%   \end{center}
%%   \caption{Voronoi Map construction in dimension 2.\label{fig:2Dbreu}}
%% \end{figure}

Algorithm \ref{alg:voro} describe the 1D processes to perform on each
row, columns and higher dimension spans. In this process, metric
information are embedded in the following two predicates (see Fig.~\ref{fig:predicates}):
\textsc{Closest}$(a, b, c)$: given three points $a,b,c\in\Z^n$ this
predicate returns true if $d(a,b) < d(a,c)$. \textsc{HiddenBy}$(a, b,
c, S)$: given three points $a,b,c\in\Z^n$ such that $a_i<b_i<c_i$ and
a 1D image span $S^i$, this predicates returns true if there is no
$s\in S$ such that
    \begin{equation}
      d(b,s) < d(a,s)\text{ and }  d(b,s) < d(c,s)\,.
    \end{equation}
\begin{algorithm}[H]\footnotesize
  \KwData{Input binary map $I_X$ if $i=1$ or  partial Voronoi map $\Pi_X$ obtained for
    dimension $1<j<i$, and a 1D segment $S$ with points
    $\{s_1,\ldots,s_{N_i}\}$ sorted by their $i^{th}$ coordinate.}
  \KwResult{Updated partial Voronoi map $\Pi_X$ along $S$.}
\If{ i = 1}
{ $k=0$\;
\ForEach{point $s_i$ in $S$}{
    \If{$I_X(s_i) = 1$}
     {
       $L_S[k] = s_i$\;
       $k++$\;
     }
}
}
\Else{
 ${L}_S[0]=\Pi_X(s_1)$\;
  ${L}_S[1]=\Pi_X(s_2)$\;
  $k=2$ , $l=3$\;
  \While{$l\leq N_i$}{
    $w = \Pi_X(s_l)$\;
    \While{$k\geq 2$ and \textsc{HiddenBy}(${L}_S[k-1],{L}_S[k],w, S$)}
          {
         $ k--$\;
        }
        $k++$ , $l++$\;
        ${L}_S[k]=w$\;
  }
}
  \ForEach{point  $s$ in $S$ by increasing $i^{th}$ coordinate}
  {
    \While{($k < |L_S|$) and not(\textsc{Closest}($s$, $L_S[k]$, $L_S[k+1]$))}
          {
            $k++$\;
          }
    $\Pi_X[s] = L_S[k]$\;
  }

  \caption{Voronoi map pruning along 1D segment $S$ along the $i^{th}$
    dimension.\label{alg:voro}}
\end{algorithm}
\begin{figure}
  \begin{center}
    \subfigure[]{ \begin{overpic}[width=2cm]{figs/hiddenByTrue}
    \put(26,5){$a$}
    \put(0,35){$b$}
    \put(25,75){$c$}
    \put(55,-15){$S$}
    \end{overpic}}
    \subfigure[]{\begin{overpic}[width=2cm]{figs/hiddenByFalse}
    \put(37,8){$a$}
    \put(23,37){$b$}
    \put(30,72){$c$}
    \put(55,-15){$S$}
    \end{overpic}}~~~~~~~~~~
    \subfigure[]{\begin{overpic}[width=1.1cm]{figs/closest}
    \put(-5,52){$a$}
    \put(2,78){$b$}
    \put(47,55){$c$}
    \put(38,-15){$S$}
      \end{overpic}
    }
  \end{center}
  \caption{xxxxxxxxxxx}
  \label{fig:predicates}
\end{figure}

In other words, \textsc{HiddenBy} returns true if and only if the
Voronoi cell of sites $a$ and $c$ \emph{hide} the Voronoi cell of $b$
along $S$ (\emph{e.g.} $A$ and $E$ \emph{hide} $C$ and $B$ on $S$ in
Fig.~\ref{fig:predicates}).  For $L_1$, $L_2$ and $L_\infty$ metrics,
\textsc{Closest} and \textsc{HiddenBy} predicates can be computed in
$O(1)$ \cite{Breu1995,Hirata1996,Meijster2000}. Hence, Algorithm
\ref{alg:voro} is in $O(N_i)$ for dimension $i$, leading to an overall
computational time for the Voronoi map and distance transformation
problem in $O(n\cdot N^n)$ (if we assume that $\forall i\in[1\ldots
  n], N_i=N$).

>>>>

\documentclass{llncs}
\usepackage{multirow}
\usepackage{amsmath,amssymb}
\usepackage{url}
\usepackage{overpic}
\usepackage{enumerate}
\usepackage{graphicx}        % standard LaTeX graphics tool
\usepackage{tikz}        % standard LaTeX graphics tool
\usepackage{subfigure}                                 % authors: subfigures
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}   % authors: last version of algorithm display
\usepackage{todonotes}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}

@@

\title{2D Subquadratic Separable Distance Transformation for
  Path-Based Norms\thanks{This work has been mainly funded by
    ANR-11-BS02-009 and ANR-11-IDEX-0007-02 PALSE/2013/21 research
    grants.}}

%% \title{Generic Separable Algorithms for Distance Transformation: A
%%   Subquadratic Approach for Path Based Norms\thanks{This work has been
%%     mainly funded by ANR-11-BS02-009 and ANR-11-IDEX-0007-02
%%     PALSE/2013/21 research grants.}}

@@

\begin{abstract}\sloppy
In many applications, separable algorithms have demonstrated their
efficiency to perform high performance volumetric computations, such
as distance transformation or medial axis extraction. In the
literature, several authors have discussed about the conditions on the
metric to be considered in a separable approach.  In this article, we
present generic separable algorithms to efficiently compute Voronoi
maps and distance transformations for a large class of
metrics. Focusing to path based norms (chamfer masks, neighborhood
sequences, ...), we detail a subquadratic algorithm to compute such
volumetric transformations in dimension 2. More precisely, we describe
a $O(\log^2{m}\cdot N^2)$ algorithm in dimension 2 for shapes in a
$N\times N$ domain with chamfer norm of size $m$.

\keywords{Digital Geometry, Distance Transformation, Path-based Norms}
\end{abstract}

@@

Since early works on digital geometry, distance transformation plays
an important role in many applications
\cite{Rosenfeld1966,Rosenfeld1968}. Given a finite input shape
$X\subset \Z^n$, the distance transformation labels each point in $X$
with the distance to its closest point in $\Z^n \setminus X$ (labeling
each point by the closest background point leads to Voronoi maps, see
below). Since such characterization is parametrized by a distance
function, many works have been done to address this distance
transformation problem with a trade-off between algorithmic
performances and the \emph{accuracy} of the digital distance function
with respect to the Euclidean one.  Hence, authors have considered
distances based on chamfer masks
\cite{Rosenfeld1968,borgefors,fouard:ivc:2005} or sequences of chamfer
masks \cite{ROSEN_66,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13};
the vector displacement based Euclidean distance
\cite{danielson,ragnemalm}; the Voronoi diagram based Euclidean
distance \cite{BreuEtAl95,maurer_pami} or the square of the Euclidean
distance \cite{Hirata,roerdnik}.  For Euclidean or $L_p$ metrics,
separable volumetric computations have demonstrated to be very
efficient: optimal $O(n\cdot N^n)$ time algorithms for shapes in $N^n$
domains, optimal multi-thread/GPU implementation\ldots (please refer
to \cite{dcoeurjo_ChapDTWADGMM} for a discussion). For path-based
metrics (chamfer mask, -weighted- neighborhood sequences,\ldots), two
main techniques exist to compute the distance transformation. The
first one considers a weighted graph formulation of the problem, and
shortest path in graphs using Dijkstra-like algorithm can be
considered. If $m$ denotes the size of the chamfer mask, computational
cost could be in $O(m\cdot N^n)$ using a cyclic bucket data structure
\cite{verwer_uniform}. Another approach consists in a raster scan of
the domain: first the chamfer mask is decomposed into two disjoint
sub-masks. Then the domain grid points are scanned in a given order
(consistent with the sub-masks construction) and a local minimum
computation is performed and then propagated
\cite{ROSEN_66,borgefors}. Scanning the domain twice for each
sub-masks leads to the distance transformation. Again, we end up with
a $O(m\cdot N^n)$ computational cost. This complexity is said to be
quadratic in $m$ and $N$ since for arbitrary large image $N$, even if
$m \ll N$, $m$ has to be in $O(N^n)$ to have accurate asymptotic DT.
Please note that Dijkstra's graph approach allows us to defined
constrained distance transformation (\emph{i.e.} geodesic metric),
both separable approaches and raster-scan for path-based metrics are
only dedicated to compact convex domains (usually hyper-rectangular)
distance transformation.

\textbf{Contributions} In this article, we first describe the generic
framework for separable distance transformation and conditions on the
metric to be consistent with this model. Then, we describe
subquadratic algorithms in dimension 2 to compute error-free distance
transformation and Voronoi map for chamfer norms and other path-based
metrics. Overall computational costs can be summarized as follows:
%\begin{table}
%  \caption{Computational cost summary for separable Voronoi map
%    computation on $N^n$ domains ($m$ being the size of the chamfer
%    norm and $f$ the number of row in a H-representation, see below).}
 % \label{tab:final}
  \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      Metric &\textsc{Closest}& \textsc{HiddenBy} & Sep. Voronoi Map & Reference\\
      \hline
      $L_2$ & $O(1)$ &  $O(1)$ & $\Theta(n\cdot N^n)$ & \cite{Hirata1996}\\
      $L_\infty$ & $O(1)$ & $O(1)$ &  $\Theta(n\cdot N^n)$ & \cite{roerdnik}\\
      $L_1$ & $O(1)$ &   $O(1)$ &$\Theta(n\cdot N^n)$& \cite{roerdnik}\\
      $L_p$  (exact pred.) & $O(\log{p})$ &
      $O(\log{p}\cdot\log{N})$ & $O(n\cdot
      N^n\cdot\log{p}\cdot\log{N})$& Lem.~\ref{lem:generic},  \cite{dgtal}\\
      $L_p$  (inexact pred.) & $O(1)$ &
      $O(\log{N})$ & $O(n\cdot
      N^n\cdot\log{N})$& Lem.~\ref{lem:generic},  \cite{dgtal}\\
      2D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
      &$O(\log^2{m}\cdot N^2)$& Theorem \ref{them}\\
  %    3D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
  %    &$O(\log^2{m}\cdot N^3)$& Theorem \ref{them2}\\
  %    nD Chamfer norm &  $O(n\cdot\log{f})$ &$O(n\cdot log^2{f}\cdot\log{N})$
  %    & $O(n^2\cdot N^n\cdot \log{f}\cdot\log{N})$& Corollary \ref{coro2}\\
      2D Neig. seq. norm  & $O(\log{f})$&
      $O(\log^2{f})$& $O(\log^2{}f\cdot N^2)$& Th.~\ref{them} and  \cite{DBLP:conf/dgci/NormandSE13}\\
      \hline
    \end{tabular}
  \end{center}
%\end{table}

\section{Preliminaries}
\label{sec:preliminaries}
%% Let us first define some elementary concepts and objects.
%% \begin{definition}[Metrics and Metric Space]
%%   \label{def:distance}
%%   A metric is a map $d$ from $E$ to a sub-group $F$ of $\R$ such that
%%   $\forall a,b,c\in E$,
%%   \begin{align}
%%     (\text{non-negative})\quad & d(a,b)\geq 0\\
%%     (\text{identity of indiscernibles}) \quad&  d(a,b)= 0
%%     \Leftrightarrow a=b\\
%%     (\text{symmetry})\quad &  d(a,b)=d(b,a)\\
%%     (\text{triangular inequality})\quad &   d(a,c) \leq d(a,b) + d(b,c)
%%   \end{align}
%% $(E, F, d)$ is called a metric space.
%% \end{definition}
%% If property $(4)$ is omitted, we are defining a distance function
%% instead of a metric.
\begin{definition}[Norm and metric induced by a norm]
  \label{def:norm}
  Given a vector space EV, a norm is a map $g$ from  $EV$ to a sub-group
  $F$ of $\R$ such that $\forall \vec{x},\vec{y}\in EV$,
  \begin{align}
    (\text{non-negative})\quad & g(\vec{x})\geq 0\\
    (\text{identity of indiscernibles}) \quad&  g(\vec{x})= 0
    \Leftrightarrow \vec{x}=\vec{0}\\
    (\text{triangular inequality})\quad &   g(\vec{x}+\vec{y}) \leq
    g(\vec{x}) + g(\vec{y})\\
    (\text{homogeneity})\quad &  \forall \lambda\in\R, \quad
    g(\lambda\cdot\vec{x}) = |\lambda|\cdot g(\cdot\vec{x})
  \end{align}
$d(a,b) \EqDef g(b-a)$ is the metric induced by the
  norm $g$.  $(E, F, d)$ is called a  metric space if $d:~E\rightarrow
  F$.
\end{definition}
Note that the above definition can be extended from vector spaces to
\emph{modules} on a commutative ring ($\Z^n$ being a module on $\Z$
but not a vector space) \cite{Thiel_hdr}. Path-based approaches
(chamfer masks, -weighted- neighborhood sequences...)  aim at defining
metrics induced by norms in metric spaces $(\Z^n, \Z , d)$.  Note that
(weighed, with $w_i\geq 0$) $L_p$ metrics
\begin{equation}
    d_{L_p} (a,b) = \left ( \sum_{k=1}^n w_k|a_k-b_k |^p \right )^{\frac{1}{p}}\;,
  \end{equation}
characterize metric spaces $(\Z^n, \R , d_{L_p})$. Rounding the \sloppy
distance function $(\Z^n,\Z, \lceil d_{L_p}\rceil)$ is a metric
space \cite{klette_book}.

\begin{definition}[Distance Transformation and Voronoi Map]
  The distance transform $DT_X$ associated with a digital metric space
  $(\Z^n,\Z,d)$ is a map  $X\rightarrow\Z$ such that, for $a\in X$
  $    DT_X(a) = \min_{b\in \Z^n\setminus X} \{d(a,b)\}$.
The Voronoi map is the map $X\rightarrow\Z^n$:
$    \Pi_X(a) = \argmin_{b\in \Z^n\setminus X} \{d(a,b)\}$.
\end{definition}
Voronoi map $\Pi_X$ corresponds to the intersection between the
continuous Voronoi diagram for the metric $d$ of points
$\Z^n\setminus$ and the lattice $\Z^n$. If a digital point $a$ belongs
to a Voronoi diagram $d-$facet ($1\leq d< n$), $a$ is equidistant to 2
or more points in $\Z^n\setminus X$ but only one is considered in
$\Pi_X(a)$ this choice has no influence on $DT_X$.
% (please refer to \cite{Couprie2007} for complete digital
%Voronoi map).
\begin{definition}[Chamfer Mask]
  A weighted vector is a pair $(\vec{v},w)$ with $\vec{v}\in\Z^n$ and
  $w\in\N^*$. A chamfer mask $\M$ is a central-symmetric set of weighted
  vectors with non-empty vectors and containing at least a basis of
  $\Z^n$.
\end{definition}
Many authors have proposed algorithmic and/or analytic approaches to
construct chamfer masks approximating the Euclidean metric and being
induced by a norm. In the following, we focus on such \emph{chamfer
  norms}.  To evaluate distances between two digital points for a
given chamfer metric, direct formulations have been proposed with
simple geometrical interpretation:
\begin{definition}[Rational ball, minimal H-representation \cite{Thiel_hdr,DBLP:journals/pr/NormandE09}]
  Given a Chamfer mask $\M$ inducing a norm, the rational ball
  associated with $\M$ is the polytope
  \begin{equation}
\label{eq:4}
    \B = conv\left\{ \frac{\vec{v}_k}{w_k};\; (\vec{v}_k,w_k)\in\M \right \}\,.
  \end{equation}
Rational ball $\B$ can also be described in a H-representation of
polytope with minimal parameter \cite{DBLP:conf/dgci/NormandSE13}: $ P
= \{ x\in\Z^n; Ax \leq y\;\}$ such that $\forall k\in[1\ldots f],
~\exists x\in P\quad A_kx=y_k.\footnote{$A_k$ being the $k^{th}$ row
  of $A$.}$ $f$ being the number of row in $A$ and the number of
facets in $\B$, and is thus related to $|\M|$.
\end{definition}
An important result for distance computation can be formalized as follows:
\begin{proposition}[Direct Distance Computation \cite{DBLP:journals/pr/NormandE09}]
  Given a chamfer mask $\M$ induced by a  norm and $(A,y)$ its minimal
  parameter H-representation, then for any $a\in\Z^n$, the chamfer
  distance of $a$ from the origin is
  \begin{equation}
    \label{eq:1}
    d_\M(O,a) =  \max_{1\leq k \leq f} \{ A_k a^T\}\,.
  \end{equation}
\end{proposition}
Among path-based digital metric, we can also introduce (weighted)
neighborhood sequences
\cite{ROSEN_66,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13}
whose objective is to have better approximation of the Euclidean
metric from sequences of elementary chamfer masks.  A key result have
been demonstrated in \cite{DBLP:conf/dgci/NormandSE13} stating that
for such distance functions, a minimal parameter polytope
representation exists and that distances can be obtained from similar
expression to (\ref{eq:1}):
\begin{equation}
  d(O,a)  =\max_{1\leq k \leq f} \{ f_k(A_k a^T)\}\,,
\end{equation}
$f_k$ being some integer sequence characterizing the neighborhood
sequence metric. In the following and for the sake of simplicity, we
describe our algorithms focusing on chamfer norms but similar results
can be obtained for more generic path-based metric such as
neighborhood sequences.
\section{Separable Distance Transformation}
\label{sec:separ-dist-transf}
\subsection{Voronoi Map from Separable Approach and Metric Conditions}
\label{sec:voronoi-map-from}
In \cite{Hirata1996,Breu1995,Meijster2000,Maurer2003},
several authors have described optimal in time and separable
techniques to compute error-free Voronoi maps or distance
transformations for $L_2$ and $L_p$ metrics. Separability means that
computations are performed dimension by dimension
%, allowing fast
%computations and multithread implementations (see
%\cite{dcoeurjo_ChapDT} for an overview).
In the following, we
consider Voronoi map approach as defined in \cite{Breu1995}: Let us
first define hyper-rectangular image $I_X:
[1..N_1]\times\ldots\times[1..N_n] \rightarrow \{0,1\}$ such that $X\subset
I_X$ and $I_X(a)= 1$ for $a\in [1..N_1]\times\ldots\times[1..N_n]$ iff $a\in
X$ ($I_X(a)=0$ otherwise). In dimension 2, we first process each row
of the input image to create independent 1D Voronoi maps along the
first dimension for the metric. Then, for each further dimension, the
partial Voronoi map $\Pi_X$ is updated using one dimensional
independent processes on image spans along the $i^{th}$ dimension.
%% \begin{figure}
%%   \begin{center}
%%      \begin{overpic}[draft=false,width=10cm]{figs/edt_dcoeurjo}
%%        \put(30,5){$Candidates_S$}
%%        \put(57,0){$S$}
%%      \end{overpic}
%%   \end{center}
%%   \caption{Voronoi Map construction in dimension 2.\label{fig:2Dbreu}}
%% \end{figure}
Algorithm \ref{alg:voro} describes the 1D processes to perform on each
row, column and higher dimensional image span\footnote{An image span
  $S$ along the $i^{th}$ direction is a vector of $N_i$ points with
  same coordinates except at $i$.} In this process, metric information
are embedded in the following two predicates (see
Fig.~\ref{fig:predicates}): \textsc{Closest}$(a, b, c)$, given three
points $a,b,c\in\Z^n$ this predicate returns true if $d(a,b) <
d(a,c)$. \textsc{HiddenBy}$(a, b, c, S)$, given three points
$a,b,c\in\Z^n$ such that $a_i<b_i<c_i$ and a 1D image span
$S$\footnote{Subscript $a_i$ denotes the $i^{th}$ coordinate of point
  $a\in\Z^n$.}, this
predicates returns true if there is no $s\in S$ such that
    \begin{equation}
      d(b,s) < d(a,s)\text{ and }  d(b,s) < d(c,s)\,.
    \end{equation}
\begin{algorithm}[H]\footnotesize
  \KwData{Input binary map $I_X$ if $i=1$ or  partial Voronoi map $\Pi_X$ obtained for
    dimensions lower than $i$, and a 1D span $S$ with points
    $\{s^1,\ldots,s^{N_i}\}$ sorted by their $i^{th}$ coordinate.}
  \KwResult{Updated partial Voronoi map $\Pi_X$ along $S$.}
\If{ i = 1}
{ $k=0$\;
\ForEach{point $s$ in $S$}{
    \If{$I_X(s) = 1$}
     {
       $L_S[k] = s$\;
       $k++$\;
\Else{
 ${L}_S[0]=\Pi_X(s^1)$\;
  ${L}_S[1]=\Pi_X(s^2)$\;
  $k=2$ , $l=3$\;
  \While{$l\leq N_i$}{
    $w = \Pi_X(s^l)$\;
    \While{$k\geq 2$ and \textsc{HiddenBy}(${L}_S[k-1],{L}_S[k],w, S$)}
          {
         $ k--$\;
        $k++$ ; $l++$\;
        ${L}_S[k]=w$\;
  \ForEach{point  $s$ in $S$ by increasing $i^{th}$ coordinate}
  {
    \While{($k < |L_S|$) and not(\textsc{Closest}($s^i$, $L_S[k]$, $L_S[k+1]$))}
          {
            $k++$\;
    $\Pi_X[s] = L_S[k]$\;

  \caption{Voronoi map construction on 1D image span $S$ along the
    $i^{th}$ dimension.\label{alg:voro}}
\end{algorithm}
\begin{figure}
  \begin{center}
    \subfigure[]{ \begin{overpic}[width=2cm]{figs/hiddenByTrue}
    \put(26,5){$a$}
    \put(0,35){$b$}
    \put(25,75){$c$}
    \put(55,-15){$S$}
    \end{overpic}}
    \subfigure[]{\begin{overpic}[width=2cm]{figs/hiddenByFalse}
    \put(37,8){$a$}
    \put(23,37){$b$}
    \put(30,72){$c$}
    \put(55,-15){$S$}
    \end{overpic}}~~~~~~~~~~
    \subfigure[]{\begin{overpic}[width=1.1cm]{figs/closest}
    \put(-5,52){$a$}
    \put(2,78){$b$}
    \put(47,55){$c$}
    \put(38,-15){$S$}
      \end{overpic}
    }
  \end{center}
  \caption{\small Geometrical predicates for Voronoi map construction:
    \textsc{HiddenBy}($a,b,c,S)$ returns false in $(a)$ and true in
    $(b)$ (straight segments correspond to Voronoi diagram
    edges). $(c)$ illustrates the \textsc{Closest}($a,b,c$) predicate
    for $c\in S$.}
  \label{fig:predicates}
\end{figure}
